### 2023-11-09 paper/content.tex
Updates from Overleaf

@@

\subsection{Prior information and uniqueness of solutions}

Working with potential field data might be very tricky once there is too much ambiguity involved during data modeling, thus to obtain unique and reliable results from data inversion it is necessary to provide as much prior information as possible.
One way to circumvent the ambiguity is to incorporate prior information about the subsurface structure, such as the position of a known source of magnetization.
Magnetic field measurements are more sensitive to changes in magnetization near the source that is causing the anomaly.
When the position of the source is known, it constrains the model to be as consistent as possible with the observed data, hence increasing the likelihood of obtaining unique solutions.
\citet{Oliveira2015Estimation} proved that the magnetization directions (Dec and Inc) recovered by the least squares estimator are sensitive to great variations in the horizontal coordinates of the center of the magnetic sources, but are practically insensitive to variations in depth.
Thus, they consider the ED method as an adequate technique to estimate the central positions that will be used as prior information for inversion.
This occurs mainly because, when well performed, the recovery of the source's horizontal coordinates is considerably accurate, while the vertical coordinate can undergo greater variation even though it still provides satisfactory results \citep{Silva20033D, Melo2013}.
This remark is also better observed in our simple synthetic sample where the estimated horizontal positions slightly deviate from the true values, which implies small misfit values in the recovered magnetic directions.
Although the estimated magnetic moment for the said sample is satisfactory, this magnetic parameter is more affected by the variations in the depth of the source, which is probably caused by ambiguities.
In summary, in order to estimate all magnetic parameters as precisely as possible the ED must be executed within a data window containing the lowest amount possible of noise since that high-frequency noise sensibility is one of the ED's main limitations.

\subsection{A critical examination of the source detection}

@@

@@

The complex synthetic data allows better observation of the strengths and limitations of the windows approach. The main strengths that can be mentioned are: 
% (i) the applied technique not only detects most of the modeled sources but also (ii) most of the recovered magnetic parameters have considerably low errors, especially in the directions, usually less than 5° (Figure~\ref{complex-synthetic-comparison}a).
% (iii) The magnetic moments obtained from well-individualized sources tend to not deviate much from the real values (Figure~\ref{complex-synthetic-comparison}b) when $R^2$ and SNR scores are considered high ($\geq 0.85$ and $\geq 5$, respectively) (Figure~\ref{complex-synthetic-comparison}c-d).
% (iv) Shallow particles grouped in clusters are usually well individualized during window selection, as well as (v) some isolated particles that produce a weak magnetic signal.
\begin{enumerate}[label=\roman*)]
  \item The applied technique detects most of the modeled sources;
  \item The majority of the recovered magnetic parameters have considerably low errors, especially in the directions, usually less than 5° (Figure~\ref{complex-synthetic-comparison}a);
  \item The magnetic moments obtained from well-individualized sources tend to not deviate much from the real values (Figure~\ref{complex-synthetic-comparison}b) when $R^2$ and SNR scores are considered high ($\geq 0.85$ and $\geq 5$, respectively) (Figure~\ref{complex-synthetic-comparison}c-d);
  \item Shallow particles grouped in clusters are usually well individualized during window selection; as well as
  \item Some isolated particles that produce a weak magnetic signal.
\end{enumerate}

It is clear from the error bars in Figure~\ref{complex-synthetic-comparison} that deep-seated sources and/or particles with small dipole moments generate worse results, essentially because they will produce weaker anomalous fields in the magnetic maps.
Note nonetheless, that even sources with small dipole moments when close to the surface are adequately modeled by our method, because of the trade-off between signal and noise for grains near the sensor. So we can point out the major limitations observed were: 
% (i) the blob detection fails when there are sources too close, grouping them into the same window, thus causing an erroneous result both for Euler deconvolution and for the magnetic parameters.
% (ii) The very same occurs when there is a source under another, in this case, the magnetic signal is the sum of both.
% (iii) In clusters of larger and/or deeper particles, although the method individualizes them well, the magnetic signal of the neighboring particles can considerably influence the result of the inversion, especially the estimated dipole moment (cluster in the Figure~\ref{complex-synthetic-comparison}b with the highest misfit values).
% As expected, there is a direct relationship between the dipole intensity and depth with the observed errors.

\begin{enumerate}[label=\roman*)]
  \item  the blob detection fails when there are sources too close, grouping them into the same window, thus causing an erroneous result both for Euler deconvolution and for the magnetic parameters;
  \item  The very same occurs when there is a source under another, in this case, the magnetic signal is the sum of both.
  \item In clusters of larger and/or deeper particles, although the method individualizes them well, the magnetic signal of the neighboring particles can considerably influence the result of the inversion, especially the estimated dipole moment (cluster in the Figure~\ref{complex-synthetic-comparison}b with the highest misfit values). As expected, there is a direct relationship between the dipole intensity and depth with the observed errors.
\end{enumerate}

\subsection{Reliability of dipole moment approximation}
\label{dipole-reliability}

Our approach relies in the premise of assuming the magnetic anomaly within
the data window is a response of a dipolar source.
The latter is true when working
with particle signals in the SD magnetic domain state since they are uniform magnetized particles with strong dipolar anomalies.
However, \cite{Nagy2017} reported that particles within the PSD domain can record the paleomagnetic field for longer (than SD ones) periods of time, being the stabler and also with strongly non-dipolar characteristics.
Therefore the application of the proposed algorithm to natural samples should fail for those PSD particles.
\cite{CortesOrtuno2022} give important insights about the matter, they showed that PSD state particles present more accurate inversion results when considering the non-dipole components for small sample-sensor distances (\textless 1 $\mu$m), but for larger sensor distances the dipole as approximations are remarkably accurate, as the higher-order moments decay rapidly with distance and therefore have less
influence on the particle's magnetic signal.
Thus, our approach can be considered reasonable to work with both particle SD and PSD states signals.
The latter happens mainly due to the sensor height being usually greater than 5 $\mu m$, considering a particle on the immediate surface of the sample the higher-order moments are already quite attenuated, thus circumventing the prominent problems described for non-dipolar components. This is corroborated by our non-dipolar synthetic sample showing the almost complete attenuation of non-dipolar components at distances greater than 5 $\mu m$.

>>>>

\subsection{Applicability to different grain concentration scenarios}

In order to assess the algorithm's ability to work with samples featuring higher concentrations of particles, we conducted multiple scenario tests (with 15 repetitions each) by increasing the concentrations of magnetic grains. Figure XX.A depicts the percentage of particle detection ($100($detected particles / total particles$)$) by the algorithm, with the median of the distribution represented by the red line. The same figure also presents the percentage of windows whose inversion results meet the filtering criteria of R-squared greater than or equal to 0.85 and SNR greater than or equal to 5, with the median represented by the blue line.

As the number of particles increases, there is a consistent decrease in the percentage of source detection, particularly due to the spatial constraints of the algorithm, which require magnetic sources to be sparsely distributed for optimal detection. However, as shown in Figure XX.B, by calculating the angular misfit between the mean direction of the model and the mean direction of the detected windows, it is observed that for the unfiltered windows (red line), the angular misfits tend to increase with the concentration of magnetic grains. Nonetheless, when applying the filtering criteria, this misfit tends to approach zero.

This observation highlights the algorithm's ability to effectively filter out spurious detection and improve the reliability of source detection, particularly in scenarios with higher particle concentrations. These findings underscore the algorithm's robustness in handling varying sample conditions and its capacity to adapt to different levels of particle density.

@@

@@

LINK THE NEXT PARAGRAPH

\subsection{NEW TITLE} %A critical examination of the source detection

@@

The complex synthetic data allows better observation of the strengths and limitations of the windows approach. The main strengths that can be mentioned are: 
(i) the applied technique not only detects most of the modeled sources but also (ii) most of the recovered magnetic parameters have considerably low errors, especially in the directions, usually less than 5° (Figure~\ref{complex-synthetic-comparison}a).
(iii) The magnetic moments obtained from well-individualized sources tend to not deviate much from the real values (Figure~\ref{complex-synthetic-comparison}b) when $R^2$ and SNR scores are considered high ($\geq 0.85$ and $\geq 5$, respectively) (Figure~\ref{complex-synthetic-comparison}c-d).
(iv) Shallow particles grouped in clusters are usually well individualized during window selection, as well as (v) some isolated particles that produce a weak magnetic signal.
% \begin{enumerate}[label=\roman*)]
%   \item The applied technique detects most of the modeled sources;
%   \item The majority of the recovered magnetic parameters have considerably low errors, especially in the directions, usually less than 5° (Figure~\ref{complex-synthetic-comparison}a);
%   \item The magnetic moments obtained from well-individualized sources tend to not deviate much from the real values (Figure~\ref{complex-synthetic-comparison}b) when $R^2$ and SNR scores are considered high ($\geq 0.85$ and $\geq 5$, respectively) (Figure~\ref{complex-synthetic-comparison}c-d);
%   \item Shallow particles grouped in clusters are usually well individualized during window selection; as well as
%   \item Some isolated particles that produce a weak magnetic signal.
% \end{enumerate}

It is clear from the error bars in Figure~\ref{complex-synthetic-comparison} that deep-seated sources and/or particles with small dipole moments generate worse results, essentially because they will produce weaker anomalous fields in the magnetic maps.
Note nonetheless, that even sources with small dipole moments when close to the surface are adequately modeled by our method, because of the trade-off between signal and noise for grains near the sensor. So we can point out the major limitations observed were: 
(i) the blob detection fails when there are sources too close, grouping them into the same window, thus causing an erroneous result both for Euler deconvolution and for the magnetic parameters.
(ii) The very same occurs when there is a source under another, in this case, the magnetic signal is the sum of both.
(iii) In clusters of larger and/or deeper particles, although the method individualizes them well, the magnetic signal of the neighboring particles can considerably influence the result of the inversion, especially the estimated dipole moment (cluster in the Figure~\ref{complex-synthetic-comparison}b with the highest misfit values).
As expected, there is a direct relationship between the dipole intensity and depth with the observed errors.

% \begin{enumerate}[label=\roman*)]
%   \item  the blob detection fails when there are sources too close, grouping them into the same window, thus causing an erroneous result both for Euler deconvolution and for the magnetic parameters;
%   \item  The very same occurs when there is a source under another, in this case, the magnetic signal is the sum of both.
%   \item In clusters of larger and/or deeper particles, although the method individualizes them well, the magnetic signal of the neighboring particles can considerably influence the result of the inversion, especially the estimated dipole moment (cluster in the Figure~\ref{complex-synthetic-comparison}b with the highest misfit values). As expected, there is a direct relationship between the dipole intensity and depth with the observed errors.
% \end{enumerate}

\subsection{Critical assessment of source position determination} %Prior information and uniqueness of solutions

Working with potential field data might be very tricky once there is too much ambiguity involved during data modeling, thus to obtain unique and reliable results from data inversion it is necessary to provide as much prior information as possible.
One way to circumvent the ambiguity is to incorporate prior information about the subsurface structure, such as the position of a known source of magnetization.
Magnetic field measurements are more sensitive to changes in magnetization near the source that is causing the anomaly.
When the position of the source is known, it constrains the model to be as consistent as possible with the observed data, hence increasing the likelihood of obtaining unique solutions.
\citet{Oliveira2015Estimation} proved that the magnetization directions (Dec and Inc) recovered by the least squares estimator are sensitive to great variations in the horizontal coordinates of the center of the magnetic sources, but are practically insensitive to variations in depth.
Thus, they consider the ED method as an adequate technique to estimate the central positions that will be used as prior information for inversion.
This occurs mainly because, when well performed, the recovery of the source's horizontal coordinates is considerably accurate, while the vertical coordinate can undergo greater variation even though it still provides satisfactory results \citep{Silva20033D, Melo2013}.

This remark is also better observed in our simple synthetic sample where the estimated horizontal positions slightly deviate from the true values, which implies small misfit values in the recovered magnetic directions.
Although the estimated magnetic moment for the said sample is satisfactory, this magnetic parameter is more affected by the variations in the depth of the source, which is probably caused by ambiguities.
In summary, in order to estimate all magnetic parameters as precisely as possible the ED must be executed within a data window containing the lowest amount possible of noise since that high-frequency noise sensibility is one of the ED's main limitations.

\subsection{Critical assessment of dipole moment approximation} %Reliability of dipole moment approximation
\label{dipole-reliability}

Our approach relies in the premise of assuming the magnetic anomaly within
the data window is a response of a dipolar source.
The latter is true when working
with particle signals in the SD magnetic domain state since they are uniform magnetized particles with strong dipolar anomalies.
However, \cite{Nagy2017} reported that particles within the PSD domain can record the paleomagnetic field for longer (than SD ones) periods of time, being the stabler and also with strongly non-dipolar characteristics.
Therefore the application of the proposed algorithm to natural samples should fail for domains bigger than PSD state. In these instances, where sources may exhibit higher-order magnetic anomalies influence such as quadrupole or octupole \citep[e.g.,][]{CortesOrtuno2021, Koster2023lava}. In such cases, the dipole model fitting process may not yield satisfactory results. Consequently, the presence of higher-order anomalies within the magnetic sources can impact the accuracy of our inversion results.

\cite{CortesOrtuno2022} give important insights about the matter, they showed that PSD state particles present more accurate inversion results when considering the non-dipole components for small sample-sensor distances (\textless 1 $\mu$m), but for larger sensor distances the dipole as approximations are remarkably accurate, as the higher-order moments decay rapidly with distance and therefore have less
influence on the particle's magnetic signal.
Thus, our approach can be considered reasonable to work with both particle SD and PSD states signals.
The latter happens mainly due to the sensor height being usually greater than 5 $\mu m$, considering a particle on the immediate surface of the sample the higher-order moments are already quite attenuated, thus circumventing the prominent problems described for non-dipolar components. This is corroborated by our non-dipolar synthetic sample showing the almost complete attenuation of non-dipolar components at distances greater than 5 $\mu m$.
